Subsetting plant species list by fruiting species

### 2024-05-30 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
add subset of trait and occurrence data by fruitng species list and write resulting data to csv

---
title: "Subsetting plant species list by fruiting species"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant occurrence and trait data by fruiting plant species."
data input: "ADD"
data output: "ADD"
date: "2024-05-29"
output: html_document
---

@@

```{r}
traits %>%
  count(TraitName)
```

```{r}
traits %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

```{r}
# Group by TraitName and summarize the unique TraitValues
unique_trait_values <- traits %>%
  group_by(TraitName) %>%
  summarize(UniqueValues = list(unique(TraitValue)))%>%
  unnest(UniqueValues)
```

@@

# Fruit type
```{r}
traits_fruitType <- traits %>%
  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue

@@

```{r}
# standardize trait values
traits_fruitType <- traits_fruitType %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
    grepl('capsule|Capsule|Capsule ', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
    grepl('follicle|Follicle', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
  )) %>%
  filter(!is.na(TraitValue))
```

@@

```{r}
summary(averagetraits_fruitType)
```

>>>>

---
title: "Subsetting plant species list by fruiting species"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant occurrence and trait data by fruiting plant species."
data input: "TropicalAndes_GBIF_plant_occ_harmonized.csv, TropicalAndes_TRY_plant_traits_harmonized.csv,
             TropicalAndes_BIEN_plant_traits_harmonized.csv, TropicalAndes_GIFT_plant_traits_harmonized.csv"
data output: "ADD"
date: "2024-05-29"
output: html_document
---

@@

```{r}
traits_clean %>%
  count(TraitName)
```

```{r}
traits_clean %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

@@

# Fruit type
```{r}
traits_fruitType <- traits_clean %>%
  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue

@@

```{r}
# standardize trait values
traits_fruitType <- traits_fruitType %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
    grepl('capsule|Capsule|Capsule ', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
    grepl('drupe|Drupe|Drupe |Drupaceous', OrigTraitValue, ignore.case = TRUE) ~ "drupe",
    grepl('aggregate berries', OrigTraitValue, ignore.case = TRUE) ~ "aggregate berries",
    grepl('aggregate drupelets', OrigTraitValue, ignore.case = TRUE) ~ "aggregate drupelets",
    grepl('fleshy', OrigTraitValue, ignore.case = TRUE) ~ "fleshy",
    grepl('pome', OrigTraitValue, ignore.case = TRUE) ~ "pome",
    grepl('Squash', OrigTraitValue, ignore.case = TRUE) ~ "squash",
  )) %>%
  filter(!is.na(TraitValue))
```

@@

```{r}
averagetraits_fruitType$TraitValue <- as.factor(averagetraits_fruitType$TraitValue)
summary(averagetraits_fruitType)
```

```{r}
fruiting_species <- averagetraits_fruitType$Accepted_species
```

# Subset trait data and occurrence data by fruiting species list
```{r}
nrow(plant_occ)
plant_occ_subset <- plant_occ %>%
  filter(Accepted_species %in% fruiting_species)
nrow(plant_occ_subset)
```

```{r}
nrow(traits_clean)
traits_subset <- traits_clean%>%
  filter(Accepted_species %in% fruiting_species)
nrow(traits_subset)
```

```{r}
traits_subset %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

# Write data to csv
```{r}
write.csv(plant_occ_subset, file.path(output_path, "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv"))
write.csv(traits_subset, file.path(output_path, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
```
